---
title: "HarvardX - Probability"
author: "Hampus Lennaárd"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style type="text/css">
  body{
  font-size: 12pt;
}
</style>

```{r library, echo=TRUE, warning=FALSE, message=FALSE}
library(tidyverse)
```

```{r set-seed, echo=TRUE, warning=FALSE, message=FALSE}
set.seed(1992, sample.kind = "Rounding")
```

## 1.0 The HardvardX Probability Course

This Markdown file was written while I was progressing through Harvard's micromaster program Data Science. Specifically, this document relates to the course on probability. The document contains both exercises and code that was provided in discussion on course material. 

### A note on seeds

While normally setting a seed is done using set.seed(1992), this course specifically instructs its participants to use the old method for setting seeds. By adding sample.kind = "Rounding" to the set.seed() argument, we revert to this older method (Version: R < 3.6)

### Notations

In the course there are several notations used which are outlined below: 

- Pr(A) : The probability of event "A" occuring \\
- X >= 6 : The event of (ex) a person, being over 6ft tall, used in relation to continous variables \\

## 2.0 Discrete Probability

Discrete probability is the subfield of probability that is associated with categorical data, or events where outcomes are not attributable to random variables. As an example, card games and other games of chance are based on discrete probability. The probability of picking a *King* from a fresh deck of cards is 4/52, and this always holds because in a fresh deck of cards there are always 4 kings and a total of 52 cards.

When we know the relative frequency of different categories within a sample, the distribution of outcomes is easily defined. As illustrated in the example above, the probability of picking a *King* from a fresh deck of cards = 0.07%. 

## 2.1 Monte Carlo Simulation

This type of simulation is used to perform experiments where we for example pick a red bean from an urn of 2 red cards and 3 blue beans. By using random number generators, we can easily perform these type of experiments. In the code below, an urn is defined using the rep() function which repeats *red* and *blue* 2 and 3 times respectivley.

By using the sample() function on beads and applying sample() once, we pick a random value from the defined vector. A random outcome is produced. However, because the argument *replacement* in sample() is not set, we limit the amount of times this "experiment" can be replicated to a total of 5 times, since there are in total 5 beans. By adding *replacement* we create a very basic example of a monte carlo simulation, allowing us to replicate the experiment an "infinite" (practically equivalent) number of times.

Another way of conducting a monte carlo simulation is to apply the replicate() function to the sample(beads, 1) call. This allows us to repliacte the experiment with replacement a set number of times, in the example we define "B" as 10K, indicating we want to replicate the experiments 10.000 times. 

```{r monte_carlo_no_replacement, echo=TRUE, message=FALSE, warning=FALSE}

# Here we define "beads" where we instruct R to create a vector of red and blue beans, indicating that we want 2 red and 3 blue beans. 
beads <- rep(c("red", "blue"), times = c(2,3))

sample(beads, 1)

B <- 10000

events <- replicate(B, sample(beads, 1))

tab <- table(events)

prop.table(tab)

```
